---
format: gfm
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# b32

<!-- badges: start -->
[![extendr](https://img.shields.io/badge/extendr-^0.8.1-276DC2)](https://extendr.rs/extendr/extendr_api/)
<!-- badges: end -->

<!-- badges: start -->
[![R-CMD-check](https://github.com/extendr/b32/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/extendr/b32/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

b32 provides fast, vectorized base32 encoding and decoding with zero dependencies powered by rust 🦀

::: callout-tip
Need base64 instead? Check out [`{b64}`](https://github.com/extendr/b64).
:::

## Installation

You can install the development version of b32 like so:

```r
pak::pak("extendr/b32")
```

## Example

Encode to base32 using `encode()`:

```{r example}
library(b32)

encoded <- encode(c("Hello", "from", "extendr"))
encoded
```

Decode using `decode_as_string()` to get back your original string:

```{r}
decode_as_string(encoded)
```

You can also decode to raw vectors with `decode()` which returns a `"blob"` object. The [`blob`](https://github.com/tidyverse/blob/) package is only a suggested dependency, but if you attach it, its print method will be used:

```{r}
library(blob)

decode(encoded)
```

